---
title: "Lending_Club"
author: "Santiago Torres"
date: "December 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:\\Users\\TORRES_SANT\\Documents\\Programming\\Data\\lending_club_analysis")
```

# Read data

```{r}
defaults <- read.csv('cleaned_up.csv')
summary(defaults$loan_status)
```
I only care if a loan has been charged off or defaulted so i will drop the values that arent current, fully paid, charged off, or default

```{r}
keep <- defaults$loan_status %in% c('Charged Off', 'Current', 'Default', 'Fully Paid')
ndefaults <- defaults[keep, ]
binaryResponse <- mapply(function(x) { 
  x %in% c('Current','Fully Paid')
  }, ndefaults$loan_status)
ndefaults$loan_status <- binaryResponse
summary(ndefaults$loan_status)
```

These are all individual accounts that meet Lending Club's policies with initial listing status - F not on a payment plan

I want to predict False in loan_status

## Response

`loan_status` - Categorical response with 2 levels:

    {
    True: Loan is current or fully paid
    False: Loan is charged off or in default
    }

## Predictors

```{r}
names(ndefaults)
length(names(ndefaults))
```

We still have a lot of predictors here lets try to consolidate some of them

### Quantitative

`loan_amt`
`int_rate`
`installment`
`annual_inc`
`dti`
`delinq_2yrs`
`earliest_cr_line` (length of time)
`fico_range_low`
`fico_range_high` (take avg)
`inq_last_6mths`
`mths_since_last_delinq`
`mths_since_last_record`
`open_acc`
`pub_rec`
`revol_util`
`total_acc`

We only need one predictor representing the loan amount to avoid multicollinearity 
```{r}
ndefaults$funded_amnt <- NULL
ndefaults$funded_amnt_inv <- NULL
```

Drop `revol_bal` because `revol_util` is a better representation of usage (ratio)
```{r}
ndefaults$revol_bal <- NULL
```

We will average FICO scores when they got the loan

### Categorical

`term`
{
36 months
60 months
}
make sure to factor categorical variables
```{r}
ndefaults$term <- factor(ndefaults$term)
```

`emp_title` is outside the scope of our investigation since it would require text analysis; `desc` and `title` are also out of scope.

```{r}
ndefaults$emp_title <- NULL
ndefaults$desc <- NULL
ndefaults$title <- NULL
```

`grade`
`subgrade`
* `emp_length`

* `home_ownership`
  + Rent; Mortgage; Own

`verification_status`
`issue_d`
`purpose`
`zip_code` is a more granular version of `addr_state`
```{r}
ndefaults$zip_code <- NULL
```
`addr_state`

Whittling down predictors gets harder so it is time to begin some analysis

Quantitative
`loan_amt`
`int_rate`
`installment`
`annual_inc`

```{r}
pairs(~ loan_amnt + int_rate + installment + annual_inc, data=ndefaults, main="Scatterplot Matrix of Defaults with Quantitative Variables")
```

`dti`
`delinq_2yrs`
`earliest_cr_line` (length of time)
`fico_range_low`
`fico_range_high` (take avg)
`inq_last_6mths`
`mths_since_last_delinq`
`mths_since_last_record`
`open_acc`
`pub_rec`
`revol_util`
`total_acc`

```{r}
hist(ndefaults$loan_amnt, main="Bill Amounts")
```

#### KNN

```{r eval=FALSE}
library(DMwR)
dknn <- ndefaults only quantile

# standardize quant

dknn[ ,1:x] <- scale(dknn[ ,1:x])

# partition

idxs = sample(1:nrow(defaults), as.integer(0.8*nrow(defaults)))

traindefaults <- defaults[idxs, ]
testdefaults <- defaults[-idxs, ]

nn7 <- kNN(loan_status ~ quants, traindefaults, testdefaults, norm=FALSE, k=7)

table(testdefaults[ ,"DEFAULT"], nn7)
```

#### Naive Bayes

#### Classification Tree

```{r eval=FALSE}
library(rpart)
fit <- rpart(loan_status ~., method="class", data=ndefaults, control=rpart.control(minsplit=1, xval=10))
printcp(fit)

plot(fit, uniform=TRUE, main="UnPruned Classification Tree for defaults")
```

Prune
```{r eval=FALSE}
pfit<- prune(fit, cp=fit$cptable[which.min(fit$cptable[,"xerror"]),"CP"])
printcp(pfit) # display the results

plot(pfit, uniform=TRUE, main="Pruned Classification Tree for defaults")
```

#### Logistic Regression

```{r}
library(MASS)
library(bestglm)

train <- ndefaults[1:(floor(.80 * nrow(ndefaults))), ]
test <- ndefaults[(ceiling(.80 * nrow(ndefaults))):nrow(ndefaults), ]
#Stepwise Regression (AIC Criterion)
# fit <- glm(loan_status ~ .,data=train,family=binomial(link='logit')) # gotta fix a lot of the categorical predictors  and some quants are read as cat
fit <- glm(loan_status ~ loan_amnt + term + annual_inc + home_ownership + verification_status + dti + delinq_2yrs + open_acc + total_acc,data=train,family=binomial(link='logit'))
summary(fit)
# step <- stepAIC(fit, direction="backward")
# step$anova # display results
```

```{r eval=FALSE}
#All Subset Selection (AIC criterion)
out <- bestglm(ndefaults, family = binomial, IC = "AIC", method = "exhaustive")
#show results for the top model 
out
```